---
title: "Assignment_4"
author: "YOUR NAME HERE"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For assignment 4, we are going to be using the built in mtcars dataset in R.

## Question 1 (.5pts) using the `data()` function, load the mtcars dataset

```{r}
```

## Question 2 (1 pts)

Write an if else statement inside a for loop that prints out if a car has high or low horsepower. Assume that horsepower (hp) greater than 150 is high.

```{r}

```

## Question 3 (.5 pt)

Using %in% and `rownames()` check if "Mazda RX4" is in the mtcars dataset.

```{r}

```

## Question 4 (1pts)

Using %in% and the table() function, see how many cars in the mtcars dataset have 4 gears (gear).

```{r}
```

## Question 5 (2 pts)

Write a while loop that calculates the cumulative sum of mpg until the cumulative sum exceeds 200.

```{r}

```

## Question 6 (3 pts)

Write a for loop and combine it with an if else statement to print the names of cars that have above-average miles per gallon (mpg).

```{r}
# avg_mpg <- mean(mtcars$mpg) # to start, I'll give you the average mpg
```

## Question 7 (2 pts)

In your own words, describe what a for loop does:

Once you are done, knit the document to html using the knit button at the top of R stuido. Please submit both your .rmd file as well as your knitted .html file to Canvas
