---
title: "Coding Lab 6"
author: "Hannah Overbye-Thompson"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Introduction and Announcements**

# **Other Programming Things You Need to Know**

## Conditional Statements

Conditional statements are a fundamental aspect of programming that allow you to control the flow of execution based on certain conditions. In R, the primary tool for creating conditional logic is the `if...else` statement.

```{r, eval = FALSE}
if (condition) {
  # Code to execute if condition is true
} else {
  # Code to execute if condition is false
}
```

-   **Condition**: This is a logical expression that evaluates to either `TRUE` or `FALSE`.

-   **Code to execute if condition is true**: This block of code runs if the condition is `TRUE`.

-   **Code to execute if condition is false**: This block of code runs if the condition is `FALSE`.

For example:

```{r}
x <- 5

if (x > 0) {
  print("x is positive")
} else {
  print("x is not positive")
}
```

### Using `else if` for Multiple Conditions

When you have multiple conditions to check, you can use the `else if` statement. The syntax is:

```{r, eval = FALSE}
if (condition1) {
  # Code to execute if condition1 is true
} else if (condition2) {
  # Code to execute if condition2 is true
} else {
  # Code to execute if none of the conditions are true
}
```

```{r}

if (x > 0) {
  print("x is positive")
} else if (x < 0) {
  print("x is negative")
} else {
  print("x is zero")
}
```

### Vectorized `ifelse` Function

```{r}
x <- c(-1, 0, 1, 2)

result <- ifelse(x > 0, "positive", "non-positive")
print(result)
```

## While Loops

In R, the `while` loop is used to repeat a block of code as long as a specified condition is true.

```{r, eval = FALSE}
while (condition) {
  # Code to execute repeatedly as long as the condition is true
}
```

-   **condition**: This is a logical expression that is evaluated before each iteration of the loop. The loop continues as long as this condition is `TRUE`.

-   **Code to execute repeatedly**: This block of code runs repeatedly as long as the condition remains `TRUE`.

```{r}
x <- 1

while (x <= 5) {
  print(x)
  x <- x + 1
}
```

### Break Condition

Sometimes you may want to exit the loop before the condition becomes `FALSE`. You can use the `break` statement to achieve this:

```{r}
x <- 1

while (x <= 10) {
  if (x == 6) {
    break
  }
  print(x)
  x <- x + 1
}
```

## For Loops

In R, `for` loops are commonly used for tasks such as iterating over vectors or data frames.

```{r, eval = FALSE}
for (variable in sequence) {
  # Code to execute for each element in the sequence
}
```

-   **variable**: A temporary variable that takes the value of each element in the sequence during each iteration.

-   **sequence**: A vector, list, or other iterable structure that contains the elements to loop over.

-   **Code to execute**: The block of code that runs for each element in the sequence.

```{r}
numbers <- c(1, 2, 3, 4, 5)

for (num in numbers) {
  print(num)
}
```

```{r}
fruits <- list("apple", "banana", "cherry", "banana")

for (fruit in fruits) {
  print(fruit)
}
```

## The Paste Function

The `paste()` function allows you to combine multiple strings (characters!) into one, with optional separators between the elements.

```{r, eval = FALSE}
paste(..., sep = " ", collapse = NULL)
```

-   **\...**: The strings or vectors to be concatenated.

-   **sep**: A character string to separate the terms (default is a space `" "`).

-   **collapse**: An optional string to separate the results when the input is a vector (default is `NULL`).

```{r}
str1 <- "Hello"
str2 <- "world"
result <- paste(str1, str2, sep = ", ")
print(result)
```

## `%in%` Operator 

%in% is used to check if elements of one vector are present in another vector.

```{r}
x <- 3
y <- c(1, 2, 3, 4, 5)

y %in% x
```

### **Practice Questions**

For these practice questions we will be using the `iris` dataset again. First, we will need to load the data

```{r}
data("iris")
```

**Question 1**

Write an if else statement inside a for loop that prints out if an iris flower has a long or short sepal length. Assume a sepal length greater than 5 is long

```{r}
for (i in iris$Sepal.Length) { 
  if (i > 5) { 
    print("long")
  } else { 
    print("short")} 
    }
```

**Question 2a**

Using `%in%` Check to see if "virginica" is in iris species

```{r}
"virginica" %in% iris$Species
```

**Question 2a**

Using `%in%` and the `table()` function, see how many times "virginica" appears in iris species

```{r}
table(iris$Species %in% "virginica")
```

**Question 3**

Write a while loop that calculates the cumulative sum of `Sepal.Length` until the cumulative sum exceeds 100

```{r}
sum <- 0 
threshold <- 50 
i = 1 

while (sum < threshold) { 
  sum <- sum + iris$Sepal.Length[i]
  i = 1 + 1
  print(sum)
  }

```

**Question 4a**

Think of a situation when an if else statement would be useful. Write out that situation below:

**Question 4b**

Think of a situation when a while loop would be useful. Write out that situation below:

**Question 4c**

Think of a situation when a for loop would be useful. Write out that situation below:
