---
title: "Coding Lab 3 Statistics in R"
author: "Hannah Overbye-Thompson"
date: "2024-07-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Introduction and Announcements**

# **Learning Statistics in R**

Today we will be going over different statistical tests in R. To do this, we will be using a data set that is pre-built into R. Do do this, we need to run the `data` function and select the dataset "iris".

The Iris dataset comprises measurements of iris flowers from three different species: Setosa, Versicolor, and Virginica. Each sample consists of four features: sepal length, sepal width, petal length, and petal width. Additionally, each sample is labeled with its corresponding species.

```{r}
data("iris")
```

We also need to load in our required libraries:

```{r}
library(tidyverse) 
library(broom) 
```

## One sample t-test

```{r}
t_test_result <- t.test(iris$Sepal.Length, mu = 5.8) %>% tidy()
print(t_test_result)
# note if you want to turn off scientific notation, you can use options(scipen=999)
```

## Paired Samples t-test

he `iris` dataset does not have a naturally paired data structure. For demonstration, I create a paired dataset by duplicating and adding noise.

```{r}
iris_paired <- iris %>%
  mutate(Sepal.Length2 = Sepal.Length + rnorm(nrow(iris)))

# Paired samples t-test
paired_t_test_result <- t.test(iris_paired$Sepal.Length, iris_paired$Sepal.Length2, paired = TRUE) %>% tidy()
print(paired_t_test_result)
```

## Independent Samples t-test

```{r}
t_test_result <- t.test(Sepal.Length ~ Species, data = iris %>% filter(Species != "virginica")) %>% tidy()

print(t_test_result)
```

## Regression

```{r}
regression_model <- lm(Sepal.Length ~ Petal.Length, data = iris)
regression_result <- tidy(regression_model)
print(regression_result)
```

## Correlation

```{r}
correlation_result <- cor.test(iris$Sepal.Length, iris$Petal.Length)
print(correlation_result)
```

## ANOVA

```{r}
anova_result <- aov(Sepal.Length ~ Species, data = iris) %>% tidy()
print(anova_result)
```

## ANCOVA

```{r}
# ANCOVA comparing Sepal.Length among Species while controlling for Petal.Length
ancova_result <- lm(Sepal.Length ~ Species + Petal.Length, data = iris) %>% anova() %>% tidy()
print(ancova_result)
```

## MANOVA

```{r}

# MANOVA comparing Sepal.Length and Petal.Length among Species
manova_result <- manova(cbind(Sepal.Length, Petal.Length) ~ Species, data = iris)
summary(manova_result)
```

## Chi-Square Goodness of Fit Test

```{r}

chisq_test_result <- chisq.test(table(iris$Species)) # we need to use the table function to count the different species 
print(chisq_test_result)
```

## Chi-Square Test of Association

The `iris` dataset does not have two categorical variables. For demonstration, I created an artificial categorical variable.

```{r}

iris <- iris %>%
  mutate(Size = ifelse(Sepal.Length > median(Sepal.Length), "Large", "Small"))

# Chi-Square Test of Association
chisq_test_result <- chisq.test(table(iris$Species, iris$Size))
print(chisq_test_result)
```

## Logistic Regression

```{r}
# Create binary dependent variable
iris <- iris %>%
  mutate(Large = ifelse(Sepal.Length > median(Sepal.Length), 1, 0))

# Logistic Regression
logistic_model <- glm(Large ~ Petal.Length + Petal.Width, data = iris, family = binomial)

logistic_result <- tidy(logistic_model)
print(logistic_result)
```

### **Practice Question**

For the following practice questions, you will be using the dataset `mtcars`.

```{r}
data("mtcars")
```

**Question 1**

A good practice whenever you load a new dataset is to explore it a little bit. Select three variables (make sure one is continuous/numerical and at least one is categorical) and calculate descriptive statistics (mean, median and mode) or frequencies, if it is a categorical variable (to do this, use the function `table()` )

```{r}
# Your code here 
```

**Question 2**

Come up with three hypotheses, one for a question involving a regression, one for a question involving a t-test (you can pick which t-test) and one involving a Chi-Squared goodness of fit test. Write out your hypotheses below as well as why you chose that particular test.

H1 (regression)

H2 (t-test)

H3 (chi-squared goodness of fit)

**Question 3**

Now run the appropriate test to test your hypothesis below

```{r}
# your code here
```

**Question 4**

Write a summary of what you found for each hypotheses below:

H1

H2

H3
